Astrophysical calibration of gravitational-wave detectors

### 2015-09-28 docs/cosmic_cal.tex
I got all the way through it.

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{NS}[NS]{neutron star}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{NSBH}[NSBH]{neutron star-black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}
\acrodef{aLIGO}[aLIGO]{Advanced LIGO}
\acrodef{AdV}[AdV]{Advanced Virgo}
\acrodef{PSD}[PSD]{power spectral density}
\acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
\acrodef{ET}[ET]{Einstein Telescope}
\acrodef{XRT}[XRT]{x-ray telescope}

@@

We want to assess how well the calibration scale factor defined in Eqn.~\ref{eq:scalefactor} can be 
estimated from a single observed \ac{GW} associated with a particular \ac{sGRB}. As we {\it a 
priori} have no knowledge of the likely location and distance of such an event we have performed 
simulations of multiple events to see the expected distribution in the accuracy of the calibration 
scale factor recovery.

% Describe the signal model (Taylor F2 etc...)
We use the TaylorF2 waveform approximation (see e.g.\ \cite{2009PhRvD..80h4043B} and references 
therein) with a 3.5 post-Newtonian expansion in phase for modelling our signal (in both simulations 
and signal recovery).  For \ac{BNS} systems we use non-spinning waveforms under the assumption that 
spins will be negligible. For \ac{NSBH} system we use a spinning waveform, but in which only the 
black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
system's orbital angular momentum.

For a non-spinning system the general form of the frequency-domain polarisation amplitudes are
%
\begin{eqnarray}\label{eq:signal}
  \gws_{+}(f) &\propto \frac{1+\cos^{2}\iota}{D_{L}}
\mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\
  \gws_{\times}(f) &\propto
 \frac{\cos^{2}\iota}{D_{L}}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, 
\phi_{\mathrm{c}})}
\end{eqnarray}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$, with 
$\eta=m_{1}m_{2}/M^{2}$ and $M=m_{1}+m_{2}$, $\iota$ is the inclination angle, and $\nu=(\pi 
Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\rm{th}}$ detector is then given by
%
\begin{equation}
  \label{eq:gravsig}
   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
F_{\times}^k(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}
%
where $F_{+}$ and $F_{\times}$ are the antenna response functions which are dependent upon the 
polarisation angle $\psi$ and the sky position of the source $\alpha$ and $\delta$.

% Define the detectors and the noise epoch we will be using
For this analysis we will consider the advanced generation \ac{aLIGO} and \ac{AdV} detectors 
operating at their design sensitivity, with noise power spectral densities taken from
\cite{2013arXiv1304.0670L}. The network we consider is the two \ac{aLIGO} detectors (H1 and L1) and 
the \ac{AdV} detector (V1).

% Define the likelihood function we will be using
We consider that the measured data in any detector is defined as
%
\begin{equation}
  \tilde{d}_{k}(f)=\scf_{k}(\tilde{n}_{k}(f)+\gws_{k}(f,\btheta))
\end{equation}
%
where $\tilde{n}(f)$ is the Fourier transform of the true (in our case Gaussian) strain noise, 
$\btheta$ is the set of waveform parameters and $\scf_k$ is the calibration scale factor for 
the $k^{\rm th}$ detector, which we want to estimate.

\subsection{Method}

To assess the ability to estimate the unknown calibration scale factors we want to calculate their 
marginal posterior probability distribution functions (pdf). We use Bayes' theorem for which we 
need to define a likelihood function and prior pdfs on the signal parameters being estimated. In 
this analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to 
sample the posterior distribution. We use a Gaussian likelihood function given by
\begin{widetext}
\begin{equation}
p(\bm{d} | \btheta, \bm{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
det}}
  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
\scf_{k}\gws_{k,i}(\btheta) - \frac{1}{2}\left(\scf_k^2 \gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
\end{widetext}
where $\bm{d}$ is an array containing the data for all detectors, $\Delta f$ is the frequency 
resolution, $S_k$ is the $k^{\rm th}$ detector's measured one sided noise \ac{PSD} (and as such 
will contain the effect of the calibration scale factor), and the $i$ indices increment over 
frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to the lower 
and upper range in frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively. 
This frequency range was chosen as the vast majority of the \ac{SNR} for the inspiralling signals 
we use can be found within this range, so going to lower or higher frequencies provides very little 
additional information. The source parameters have been subsumed into a vector $\btheta$ and the 
calibration scale factors are within the vector $\bm{\scf}$.

@@

The primary reason why we can use coincident observations with a \ac{sGRB} as a check of detector 
calibration is that the \ac{sGRB} allows us to constrain the priors on various parameters that 
would normally be highly correlated with the signal amplitude. The most important of these is 
that the \ac{sGRB} can provide a very tight constraint on the source distance\footnote{For \ac{BNS} 
and \ac{NSBH} systems they will be observable only in the relatively nearby Universe ($z \sim 
0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems respectively), so the relation between the 
measured \ac{sGRB} redshift and distance is simple and our luminosity distance will be negligibly 
effected by the choice of cosmology \MP{Check this?}.}. In this analysis we therefore assume that 
the source distance is known (i.e.\ has a $\delta$-function prior). Another piece of information 
that we can use from having a coincident \ac{sGRB} is that the system is likely to be relatively 
close to face-on (and therefore nearly circularly polarised), which allows us to place prior 
constraints on the system inclination angle. Beam opening angles for \acp{sGRB} are relatively 
poorly constrained as they are based on only a few jet break detections, however 
\cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we assume 
that the beam opening angle is a half-normal distribution, for which the median is given by 
$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
$\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian prior on the inclination angle with zero 
mean and $\sigma=14.8^{\circ}$.

Our prior on the system component masses is dependent on whether we are considering the source 
being a \ac{BNS} system or a \ac{NSBH}. For the former case the prior we use is a Gaussian 
distribution for both components with means of $1.35\,\textrm{M}_{\odot}$ and standard deviations 
of $0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we use in our 
study, the chirp mass, which we see from equation~\ref{eq:signal} plays a role in the overall 
signal amplitude, is very well constrained by the system's phase evolution, so our prior could be 
expanded with minimal effect on the results. In the case of \ac{NSBH} systems we use the same 
prior as above for the neutron star, but for the black hole we use a prior based on the 
canonical masss distribution used for the rate results in \cite{2012PhRvD..85h2002A} with a mean of 
$5\,\textrm{M}_{\odot}$ and a standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass 
range of black holes could be quite different from this, but use this range as an example for this 
type of system. For the \ac{NSBH} systems we also require a prior on the black hole spin (for 
\ac{BNS} systems we have assumed that the spins are small enough that they will be negligible). We 
use a uniform prior on the normalised aligned spin magnitude between $-1$ and $1$.

The final important piece of information that we can make use of from the \ac{sGRB} observation is 
the sky position of the source, which we take to be known precisely.

For the time of coalescence we use a uniform prior of spanning $\pm0.01$\,seconds around the 
recorded time of the signal that would be returned by the detection pipeline. For the reference 
phase a uniform prior between 0 and $\pi$ is used, and for the polarisation angle a uniform prior 
between 0 and $\pi/2$ is used. As the signals we use are generally close to being circularly 
polarised (face-on) the phase and polarisation angle will be degenerate meaning the exact 
combination of them will have very little effect of the result.

Finally, we require a prior on the calibration scale factors for each detector. We make the 
assumption that in general the calibration applied to the detector data will be correct, so want to 
use a prior that is peaked at unity. We also assume that the probability density that the 
calibration scale factor is either larger or smaller than unity by an equivalent factor, e.g.\ 
$\scf=10$ or $\scf=0.1$, is the same. For the scale factor for each detector we we therefore use a 
log-normal distribution as our prior
\begin{equation}
 p(\scf|I) = \frac{1}{\scf\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{\scf} - \mu)^2}{2\sigma^2} 
\right)},
\end{equation}
where we chose $\sigma = 1.07$, which in turn means that having the mode at unity gives $\mu = 
1.15$. With these parameters the prior probability density at 0.1 and 10 is one tenth of that at 
1. Note that this prior does not give a symmetric amount of probability about unity, but given our 
simulation criterion described below in Section~\ref{sec:simulations} we find that the likelihood 
generally overwhelms the prior. As $\scf$ is a scale factor we also have the further constraint 
that it must be positive.

\subsection{Simulations}\label{sec:simulations}

To estimate how well we can constrain the calibration scaling for each detector in an advanced 
detector network we have performed simulations with injected \ac{BNS} and \ac{NSBH} signals at a 
range of distances (50\,Mpc to 500\,Mpc with 50\,Mpc increments for the \ac{BNS} signals and 
100\,Mpc to 900\,Mpc with 100\,Mpc increments for the \ac{NSBH} signals). The simulations used the 
two \ac{aLIGO} detectors (H1 and L1) and the \ac{AdV} detector (V1). The simulations were all 
performed in the frequency domain and spanned a frequency range from 20\,Hz to 400\,Hz. We assumed
all detectors were operating at their design sensitivity (as given by Figure~1 of 
\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each simulation based on 
the sensitivity curve, but scaled with the associated calibration scale factor.

% Describe the ensemble of simulations used
Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large number simulations at 
each distance value. When proposing injections for each simulation the source parameters were 
randomly drawn from the prior distributions given in Section~\ref{sec:priors}, with the sky 
position being drawn randomly from a uniform distribution on the sky, the coalescence times held 
fixed within the centre of the prior window, and with the calibration scaling factors for each 
detector drawn from a Gaussian distribution with a mean of one and standard deviation of $0.125$ 
(equivalent to a mean fractional offset of 10\%). However, in accepting a proposed injection as one 
to be analysed 
we introduced a criteria that the signal be ``detectable'', based on that used in 
\cite{2012PhRvD..85h2002A} of it having an \ac{SNR} of $\geq 5.5$ in at least two detectors (as 
these would be signals coincident with a \ac{sGRB} we do not include the often used further 
constraint that the total coherent network \ac{SNR} is greater than 12). This criteria means that 
at larger distances our population of injected sources was not uniformly distributed over the sky, 
and were generally closer to being circularly polarised.

These simulations have allowed us to assess how well on average we would be able to constrain the 
calibration scale for a given \ac{CBC}-\ac{sGRB} coincidence.

\section{Results}\label{sec:results}

A simple assumption that one could make would be that the uncertainties on the calibration scale 
factors (if they are independent of other parameters and have a roughly Gaussian probability 
distribution) should be given by $\sim 1/{\rm SNR}$ for each detector. As we will show below this 
assumption is reasonable, but small correlations do exist between parameters meaning that it does 
not completely hold.

For each of the simulated sources we have used the {\tt emcee} python \ac{MCMC} package 
\cite{2013PASP..125..306F} to perform parameter estimation over the unknown source parameters using 
the priors as discussed in Section~\ref{sec:priors}. In each case when calculating the likelihood 
we used an estimate of the noise \ac{PSD} based on the advanced detector design sensitivities 
(using those given in \cite{2013arXiv1304.0670L}), but simulated as if calculated by averaging 64 
separate noisy \ac{PSD} estimates and scaled with the same factor as applied to the injection and 
noise\footnote{We do not account for there being a potential difference between the estimated 
\ac{PSD} and the actual \ac{PSD} of the analysed section of data as described in e.g.\ 
\cite{2013PhRvD..88h4044L}. This difference would be very highly correlated with the calibration 
scale factor, so in reality our estimate of the scale factor would be a combination of the
calibration offset and the difference in the \ac{PSD}. As such our results on real data would be an 
upper limit on the calibration scale factor.}. This has provided posterior probability distributions 
on the calibration scale factors for each detector. Examples of the marginalised posterior 
probabilities for a \ac{BNS} system and a \ac{NSBH} system observed with the three detector network 
are shown in figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.

From these posterior distributions we have calculated the minimal 68\% credible region for the 
calibration scale factors for each detector (if these were Gaussian distributions this is 
equivalent to the region either side of the mean bounded by the $1\sigma$ intervals). For all the 
signals at each distance increment we have produced the distribution of the fractional half widths 
(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the true value. These 
are shown as boxplots in figures~\ref{fig:bnsresults} and \ref{fig:nsbhresults} for the \ac{BNS} 
and \ac{NSBH} systems respectively. The boxes show the extent from the lower to upper quartile of 
the values, whilst the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The black 
line within each box gives the median value and the star gives the mean value. Also shown on each 
plot as the dashed magenta line is the percentage of signals drawn from the prior distribution that 
fulfil the detection criterion.

\begin{figure*}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{bns_post_fig.pdf}
 \end{center}
 \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
 parameters of a \ac{BNS} system, including calibration scaling factors for the three detectors 
(H1, L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 
4.1 and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each 
of the detectors respectively.}
\end{figure*}

\begin{figure*}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{nsbh_post_fig.pdf}
 \end{center}
 \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for the unknown
 parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
(H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
the detectors respectively.}
\end{figure*}

\begin{figure*}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{scale_factor.pdf}
 \end{center}
 \caption{\label{fig:bnsresults} Distributions of the percentage accuracy at which the calibration 
scale factors can be determined for a three detector network if assuming \ac{BNS} systems (provided 
a coincident \ac{GRB} is observed and can yield a distance estimate). The boxplots span the lower 
to upper quartile range of the distributions, with the median value shown as a horizontal line 
within the box and the mean shown as a star. The dashed magenta line shows the percentage of source 
drawn from the prior distribution that would be detectable at each distance value.}
\end{figure*}

>>>>

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{NS}[NS]{neutron star}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{NSBH}[NSBH]{neutron star-black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}
\acrodef{aLIGO}[aLIGO]{Advanced LIGO}
\acrodef{AdV}[AdV]{Advanced Virgo}
\acrodef{PSD}[PSD]{power spectral density}
\acrodef{PDF}[PDF]{probability distribution function}
\acrodef{MCMC}[MCMC]{Markov chain Monte Carlo}
\acrodef{ET}[ET]{Einstein Telescope}
\acrodef{XRT}[XRT]{x-ray telescope}

@@

% the main aim of the analysis
The main aim of this work is to assess how well the calibration scale factor
defined in Eqn.~\ref{eq:scalefactor} can be estimated from a single observed
\ac{GW} associated with a particular \ac{sGRB}. As we {\it a priori} have
little knowledge of the likely location and distance of such an event we have
performed simulations of multiple events to see the expected distribution in
the accuracy of the calibration scale factor recovery.

% Describe the signal model (Taylor F2 etc...)
We use the TaylorF2 waveform approximation (see e.g.~\cite{2009PhRvD..80h4043B}
and references therein) with a 3.5 post-Newtonian expansion in phase for
modelling our signal (in both simulations and signal recovery).  For \ac{BNS}
systems we use non-spinning waveforms under the assumption that spins will be
negligible~\cm{ref?}. For \ac{NSBH} system we use a spinning waveform, but in which only
the black hole has non-negligible spin and that its rotational angular momentum
is aligned with the system's orbital angular momentum.

For a non-spinning system the general form of the frequency-domain polarisation
amplitudes are
%
\begin{align}\label{eq:signal} 
\gws_{+}(f) &\propto
\left(1+\cos^{2}\iota\right)D_{L}^{-1} \mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f,
\nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\ \gws_{\times}(f) &\propto
\left(\cos^{2}\iota\right)D_{L}^{-1}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu,
t_\mathrm{c}, \phi_{\mathrm{c}})} 
\end{align}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
with the symmetric mass ratio $\eta=m_{1}m_{2}/M^{2}$.  The total mass
$M=m_{1}+m_{2}$ where $m_{1}$ and $m_{2}$ are the component masses, $\iota$ is
the inclination angle, and $\nu=(\pi Mf)^{1/3}$.  The \ac{GW} strain measured
at the $k^{\rm{th}}$ detector is then given by
%
\begin{equation} \label{eq:gravsig} 
\gws^{k}(f) = \big[ F_{+}^{k}(\alpha,
\delta, \psi)\gws_{+}(f) + F_{\times}^k(\alpha, \delta,
\psi)\gws_{\times}(f)\big] 
\end{equation}
%
where $F^{k}_{+}$ and $F^{k}_{\times}$ are the antenna response functions which
are dependent upon the polarisation angle $\psi$ and the sky position of the
source $\alpha$ and $\delta$.

% Define the detectors and the noise epoch we will be using
For this analysis we will consider the advanced generation \ac{aLIGO} and
\ac{AdV} detectors operating at their design sensitivity, with noise power
spectral densities taken from \cite{2013arXiv1304.0670L}. The detector network
we consider consists of the two \ac{aLIGO} detectors (H1 and L1) and 
the \ac{AdV} detector (V1).

% Define the likelihood function we will be using
We consider that the measured data in any detector is defined as
%
\begin{equation}
  \tilde{d}_{k}(f)=\scf_{k}(\tilde{n}_{k}(f)+\gws_{k}(f,\btheta))
\end{equation}
%
where $\tilde{n}(f)$ is the Fourier transform of the true (in this case
Gaussian) strain noise, $\btheta$ is the set of waveform parameters and
$\scf_k$ is the calibration scale factor for the $k^{\rm th}$ detector, the
parameters we are interested in estimating.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Method}

% Introduce the Bayesian approach and the likelihood
To assess the ability to estimate the unknown calibration scale factors we
calculate their marginal posterior \acp{PDF}.  We use Bayes' theorem for which
we need to define a likelihood function and prior \acp{PDF} on the signal
parameters and the calibration scale factors. In this analysis we use the
\ac{MCMC} code {\tt emcee}~\cite{2013PASP..125..306F} to sample
the posterior distribution. We use a Gaussian likelihood function given by
%
\begin{widetext}
\begin{equation}
p(\bm{d} | \btheta, \bm{\scf}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm 
det}}
  \sum_{i=i_{\rm low}}^{i_{\rm high}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
\scf_{k}\gws_{k,i}(\btheta) - \frac{1}{2}\left(\scf_k^2 \gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
\end{widetext}
%
where $\bm{d}$ is an array containing the data for all detectors, $\Delta f$ is
the frequency resolution, $S_k$ is the $k^{\rm th}$ detector's estimated
one-sided noise \ac{PSD} (and as such will contain the effect of the
calibration scale factor), and the $i$ indices increment over frequency with
$i_{\rm low}$ and $i_{\rm high}$ corresponding to the lower and upper range in
frequencies used for our analysis, which were 20\,Hz to 400\,Hz respectively.
This frequency range was chosen as the vast majority of the \ac{SNR} for the
inspiralling signals we use can be found within this range, so going to lower
or higher frequencies provides very little additional information. The source
parameters have been subsumed into a vector $\btheta$ and the calibration scale
factors are within the vector $\bm{\scf}$.

@@

% mention of distance delta function prior
The primary reason why coincident observations with a \ac{sGRB} can be used as
a check of detector calibration is that the \ac{sGRB} allows us to constrain
the priors on various parameters that would normally be highly correlated with
the signal amplitude. The most important of these is that the \ac{sGRB} can
provide a very tight constraint on the source distance\footnote{For \ac{BNS}
and \ac{NSBH} systems they will be observable only in the relatively nearby
Universe ($z \sim 0.1$ and $z\sim 0.2$ for \ac{BNS} and \ac{NSBH} systems
respectively), so the relation between the measured \ac{sGRB} redshift and
distance is simple and our luminosity distance will be negligibly effected by
the choice of cosmology \MP{Check this?}.}. In this analysis we therefore
assume that the source distance is known (i.e.\ has a $\delta$-function
prior).~\cm{Should we not say a little bit more about this? A really accurate
distance from a really accurate redshift is only as good as the Hubble constant
which currently has $\sim$ few percent error?  We should also say something
about peculiar motion.}

% discuss the inclination and opening angle
Other information relevant for a coincident \ac{sGRB} is that the system is
likely to be relatively close to face-on (and therefore nearly circularly
polarised), which allows us to place prior constraints on the system
inclination angle. Beam opening angles for \acp{sGRB} are relatively poorly
constrained as they are based on only a few jet-break detections. A comparative
study of estimates and lower-limits of opening angles in 13 \acp{sGRB} given
in~\cite{2014ApJ...780..118F} provide a median opening angle value of $\sim
10^{\circ}$. We assume that the beam opening angle is a half-normal
distribution, for which the median is given by $\sqrt{2}\erf{}^{-1}(1/2)\sigma
\sim 0.67\sigma$. It follows that the median of a $\sim 10^{\circ}$ opening
angle corresponds to $\sigma \sim 14.8^{\circ}$. We therefore place a Gaussian
prior on the inclination angle with zero mean and a standard deviation
$\sigma=14.8^{\circ}$ truncated at values below zero.

% discuss the component mass and spin priors 
The prior on the system component masses is dependent on whether we are
considering the source being a \ac{BNS} system or a \ac{NSBH}. For the former
case the prior we use is a Gaussian distribution for both components with means
of $1.35\,\textrm{M}_{\odot}$ and standard deviations of
$0.13\,\textrm{M}_{\odot}$ \MP{(reference)}. In reality, for all the systems we
use in our study, the chirp mass, which we see from Eq.~\ref{eq:signal}
plays a role in the overall signal amplitude, is very well constrained by the
system's phase evolution, so our prior could be expanded with minimal effect on
the results. In the case of \ac{NSBH} systems we use the same prior as above
for the neutron star, but for the black hole we use a prior based on the
canonical masss distribution used for the rate results in
\cite{2012PhRvD..85h2002A} with a mean of $5\,\textrm{M}_{\odot}$ and a
standard deviation of $1\,\textrm{M}_{\odot}$. We note that the mass range of
black holes could be quite different from this, but use this range as an
example for this type of system. For the \ac{NSBH} systems we also require a
prior on the black hole spin (for \ac{BNS} systems we have assumed that the
spins are small enough that they will be negligible). We use a uniform prior on
the normalised aligned spin magnitude between $-1$ and $1$.

% discuss the known sky position
The final important piece of information that we can make use of from the
\ac{sGRB} observation is the sky position of the source. We assume that this is 
known precisely and has been obtained, in conjunction with the source redshift,
from the identification of the host galaxy.

% discuss the nuisance parameters
For the time of coalescence $t_{\mathrm{c}}$ we use a uniform prior spanning
$\pm0.01$\,seconds around the recorded time of the signal that would be
returned by the \ac{GW} detection pipeline. For the reference phase
$\phi_{\mathrm{c}}$ a uniform prior on the range $(0,\pi]$ is used, and for the
polarisation angle we use a uniform prior on the range $(0,\pi/2]$. As the
signals we use are generally close to being circularly polarised (face-on) the
phase and polarisation angle will be largely degenerate and their exact
combination will have little effect on the result.

% the scale factor prior
Finally, we require a prior on the calibration scale factors for each detector.
We make the assumption that in general the calibration applied to the detector
data will be correct, so want to use a prior that is peaked at $\scf_k=1$. We
also assume that the prior \ac{PDF} of the calibration scale factor is either
larger or smaller than unity by an equivalent factor, e.g.\ $p(\scf|I)=10$ is
equivelent to $p(\scf|I)=0.1$. For the scale factor in each detector we 
therefore use the following log-normal distribution as our prior
%
\begin{equation}\label{eq:scale_prior}
 p(\scf|I) = \frac{1}{\scf\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{\scf} - \mu)^2}{2\sigma^2} 
\right)},
\end{equation}
%
where we chose $\sigma = 1.07$, which together with the requirement that the
mode is at unity gives $\mu = 1.15$. With these parameters the prior
probability density at 0.1 and 10 is one tenth of that at 1. Note that this
prior does not give a symmetric amount of probability about unity, but given
our simulation criterion described in Section~\ref{sec:simulations} we find
that the likelihood generally overwhelms the prior~\cm{I'm not sure about this
sentence.  Why does the symmetry matter?}. As $\scf$ is a scale factor we also
have the further constraint that it must be positive and hence the log-normal
prior is a natural choice.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulations}\label{sec:simulations}

% a general description of the simulations 
To estimate how well the calibration scaling can be constrained for each
detector in an advanced detector network we have performed simulations with
injected \ac{BNS} and \ac{NSBH} signals. These signal injections were simulated
at a range of distances: 50\,Mpc to 500\,Mpc with 50\,Mpc increments for the
\ac{BNS} signals and 100\,Mpc to 900\,Mpc with 100\,Mpc increments for the
\ac{NSBH} signals. The simulations used the two \ac{aLIGO} detectors (H1 and
L1) and the \ac{AdV} detector (V1) and were all performed in the frequency
domain and spanned a frequency range from 20\,Hz to 400\,Hz~\cm{can you remind
me and maybe the reader of why a 400 Hz cutoff is OK?}. We assumed all
detectors were operating at their design sensitivity (as given by Figure~1 of
\cite{2013arXiv1304.0670L}). As such we added coloured Gaussian noise to each
simulation based on the sensitivity curve, but scaled with the associated
calibration scale factor.

% Describe the ensemble of simulations used
Separately for the \ac{BNS} and \ac{NSBH} systems we have performed a large
number~\cm{how many exactly/roughly?} simulations at each distance value. When
proposing injections for each simulation the source parameters were randomly
drawn from the prior distributions described in Section~\ref{sec:priors}, with
the sky position being drawn randomly from a uniform distribution on the sky,
the coalescence times held fixed within the centre of the prior window, and
with the calibration scaling factors for each detector drawn from a Gaussian
distribution with a mean of one and standard deviation of $0.125$ (equivalent
to a mean fractional offset of 10\%)~\cm{should we explicitly acknowledge that
we do not simulate from the same priors that we assume in the recovery?}.
However, in accepting a proposed injection as one to be analysed we introduced
a criteria that the signal be ``detectable'', based on that used in
\cite{2012PhRvD..85h2002A}. We therefore only analysed injections that had an
\ac{SNR} of $\geq 5.5$ in at least two detectors. Since we assume that these
signals would be coincident with a \ac{sGRB} we do not include the often used
further constraint that the total coherent network \ac{SNR} is greater than 12.
For injections close to the \ac{SNR} threshold, the application of the
threshold resulted in the preferential selection of better oriented systems.
Therefore, at larger distances the population of injected sources was not
uniformly distributed over the sky and instead favoured locations with better
antenna response.  Similarly, despite already constrained by the prior to small
inclination angles, the accepted injections show preference for sources very
close to circularly polarised.

% final clarification on the simulation distribution
These simulations have allowed us to assess how well on average we would be
able to constrain the calibration scale for a given detected \ac{CBC}-\ac{sGRB}
coincidence.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:results}

% a general and simple results statement 
A simple assumption that one could make would be that the uncertainties on the
calibration scale factors (if they are independent of other parameters and have
a roughly Gaussian probability distribution) should be given by $\sim 1/{\rm
SNR}$ for each detector. As we will show below this assumption is reasonable,
but small correlations do exist between parameters meaning that it does not
completely hold.

% a discussion of the PSD estimate
For each of the simulated sources we have used the {\tt emcee} python \ac{MCMC}
package~\cite{2013PASP..125..306F} to perform parameter estimation over the
unknown source parameters using the priors as discussed in
Section~\ref{sec:priors}. In each case when calculating the likelihood we used
an estimate of the noise \ac{PSD} based on the advanced detector design
sensitivities (using those given in~\cite{2013arXiv1304.0670L}), but simulated
as if calculated by averaging 64 separate noisy \ac{PSD} estimates and scaled
with the same factor as applied to the injection and noise~\cm{Do you estimate
the PSD using 64 segments? This sentence might be a bit confusing to the
reader.}\footnote{We do not account for there being a potential difference
between the estimated \ac{PSD} and the actual \ac{PSD} of the analysed section
of data as described in e.g.\ \cite{2013PhRvD..88h4044L}. This difference would
be very highly correlated with the calibration scale factor, so in reality our
estimate of the scale factor would be a combination of the calibration offset
and the difference in the \ac{PSD}. As such our results on real data would be
an upper limit on the calibration scale factor.~\cm{Even though we're hiding
this in a footnote is it really this bad. How big is the statistical error on
PSD estimation using 64 segments?}}. This has provided posterior probability
distributions on the calibration scale factors for each detector.  Examples of
the marginalised posterior probabilities for a \ac{BNS} system and a \ac{NSBH}
system observed with the three detector network are shown in
figures~\ref{fig:bnspost} and \ref{fig:nsbhpost} respectively.

% describe the main boxplot results
From these posterior distributions we have calculated the minimal 68\% credible
region for the calibration scale factors for each detector (if these were
Gaussian distributions this is equivalent to the region either side of the mean
bounded by the $1\sigma$ intervals). For all the signals at each distance
increment we have produced the distribution of the fractional half widths
(i.e.\ $1\sigma$) of these scale factors' confidence intervals compared to the
true value. These are shown as boxplots in figures~\ref{fig:bnsresults} and
\ref{fig:nsbhresults} for the \ac{BNS} and \ac{NSBH} systems respectively. The
boxes show the extent from the lower to upper quartile of the values, whilst
the whiskers extend from the 5$^{\rm th}$ to 95$^{\rm th}$ percentile. The
black line within each box gives the median value and the star gives the mean
value. Also shown on each plot as the dashed magenta line is the percentage of
signals drawn from the prior distribution that fulfil the detection criterion.

\begin{figure*}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{bns_post_fig.pdf}
 \end{center}
 \caption{\label{fig:bnspost} The marginalised posterior probability distributions for the unknown
 parameters of a \ac{BNS} system, including calibration scaling factors for the three detectors 
(H1, L1 and V1). The injected signal was at a distance of 250\,Mpc, had \ac{SNR} of 8.7, 10.9 and 
4.1 and percentage uncertainties in the calibration scale factors of 14\%, 11\% and 27\% for each 
of the detectors respectively.~\cm{The fonts are a bit small.  Maybe also
highlight somehow the boxes refereing to the scale factors?}}
\end{figure*}

\begin{figure*}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{nsbh_post_fig.pdf}
 \end{center}
 \caption{\label{fig:nsbhpost} The marginalised posterior probability distributions for the unknown
 parameters of a \ac{NSBH} system, including calibration scaling factors for the three detectors 
(H1, L1 and V1). The injected signal was at a distance of 450\,Mpc, had \ac{SNR} of 7.7, 10.9 and 
4.2 and percentage uncertainties in the calibration scale factors of 17\%, 14\% and 34\% for each of 
the detectors respectively.~\cm{same things about the fonts and highlighting
the scale factors.}}
\end{figure*}
